---
title: "Learn HTMLWidget"
author: "Ko, Kuei Yueh"
date: "Thursday, July 16, 2015"
output: html_document
---
```{r}
setwd("C:\\Users\\clinton\\GitHub\\CodeLearn\\LearnR")
```

# Install and Import the libraries
```{r}
#library(devtools)
# time seires charting
#install_github("rstudio/dygraphs")
# DataTables
#install_github("rstudio/DT")
# Leaflet maps
#install_github("rstudio/leaflet")
# network graphs by D3
#install.packages("networkD3")
# 3D scatterplots and globe (three.js)
#install_github("bwlewis/rthreejs")
# 3D scatterplots, lines and bars (MetricsGraphics.js)
#install_github("hrbrmstr/metricsgraphics")
# Interactive Heatmaps (d3heatmap)
#install_github("rstudio/d3heatmap")
# DigrammeR
#install_github("rich-iannone/DiagrammeR")
# epiwidgets
install_github("sdwfrost/epiwidgets")
```

Note:  
The below tutorial is from
http://www.htmlwidgets.org/
https://www.youtube.com/watch?v=dV4gtARPvu8

# Leaflet
http://rstudio.github.io/leaflet/

Leaflet is a JavaScript library for creating dynamic maps that support panning and zooming along with various annotations like markers, polygons, and popups.

Tutorial from htmlwidgets.org
```{r}
library(leaflet)
orstationc <- read.csv("C:/Users/clinton/GitHub/CodeLearn/LearnR/data/orstationc.csv")

pal <- colorQuantile("YlOrRd", NULL, n=8)
leaflet(orstationc) %>%
    addTiles %>%
    addCircleMarkers(color = ~pal(tann))
```

# Dygraphs
http://rstudio.github.io/dygraphs

Dygraphs provides rich facilities for charting time-series data in R and includes support for many interactive features including series/point highlighting, zooming, and panning.

Tutorial from htmlwidgets.org
```{r}
library(dygraphs)
dygraph(nhtemp, main="New Haven Temperatures") %>%
    dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

# MetricsGraphics
http://hrbrmstr.github.io/metricsgraphics/

MetricsGraphics enables easy creation of D3 scatterplots, line charts, and histograms.

Tutorial from htmlwidgets.org
```{r}
library(metricsgraphics)
mjs_plot(mtcars, x=wt, y=mpg) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")
```

# NetworkD3
http://christophergandrud.github.io/networkD3/

networkD3 provides tools for creating D3 JavaScript network graphs from R.

Tutorial from htmlwidgets.org
```{r}
library(networkD3)
data(MisLinks, MisNodes)
forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.4)
```

# d3Heatmap
https://github.com/rstudio/d3heatmap

Interactive heatmaps with D3 including support for row/column highlighting and zooming.

Tutorial from htmlwidgets.org
```{r}
library(d3heatmap)
d3heatmap(mtcars, scale="column", colors="Blues")
```

# DataTables
http://rstudio.github.io/DT/

DataTables displays R matrices or data frames as interactive HTML tables that support filtering, pagination, and sorting.

Tutorial from htmlwidgets.org
```{r}
library(DT)
datatable(iris, options = list(pageLength = 5))
```

# threejs
https://github.com/bwlewis/rthreejs

threejs includes a 3D scatterplot and 3D globe (you can directly manipulate the scatterplot below with the mouse).

Tutorial from htmlwidgets.org
```{r}
library(threejs)
z <- seq(-10, 10, 0.01)
x <- cos(z)
y <- sin(z)
scatterplot3js(x,y,z, color=rainbow(length(z)))
```

# DiagrammeR
http://rich-iannone.github.io/DiagrammeR/

A tool for creating diagrams and flowcharts using Graphviz and Mermaid.

Tutorial from htmlwidgets.org
```{r}
library(DiagrammeR)
grViz("
  digraph {
    layout = twopi
    node [shape = circle]
    A -> {B C D} 
  }")
```

# epiwidgets
https://github.com/sdwfrost/epiwidgets

A collection of htmlwidgets handy for epidemiology, etc.

Tutorial from the github page sdwfrost/epiwidgets 

## treewidget
This is an interactive phylogeny viewer based on the enviable skills of @spond and @stevenweaver.

However, it seems that the treewidget function is not working.
```{r}
library(epiwidgets)
nwk <- "(((EELA:0.150276,CONGERA:0.213019):0.230956,(EELB:0.263487,CONGERB:0.202633):0.246917):0.094785,((CAVEFISH:0.451027,(GOLDFISH:0.340495,ZEBRAFISH:0.390163):0.220565):0.067778,((((((NSAM:0.008113,NARG:0.014065):0.052991,SPUN:0.061003,(SMIC:0.027806,SDIA:0.015298,SXAN:0.046873):0.046977):0.009822,(NAUR:0.081298,(SSPI:0.023876,STIE:0.013652):0.058179):0.091775):0.073346,(MVIO:0.012271,MBER:0.039798):0.178835):0.147992,((BFNKILLIFISH:0.317455,(ONIL:0.029217,XCAU:0.084388):0.201166):0.055908,THORNYHEAD:0.252481):0.061905):0.157214,LAMPFISH:0.717196,((SCABBARDA:0.189684,SCABBARDB:0.362015):0.282263,((VIPERFISH:0.318217,BLACKDRAGON:0.109912):0.123642,LOOSEJAW:0.397100):0.287152):0.140663):0.206729):0.222485,(COELACANTH:0.558103,((CLAWEDFROG:0.441842,SALAMANDER:0.299607):0.135307,((CHAMELEON:0.771665,((PIGEON:0.150909,CHICKEN:0.172733):0.082163,ZEBRAFINCH:0.099172):0.272338):0.014055,((BOVINE:0.167569,DOLPHIN:0.157450):0.104783,ELEPHANT:0.166557):0.367205):0.050892):0.114731):0.295021)"
treewidget(nwk,browser=FALSE)
```

```{r}
treewidget(
"(((Crotalus_oreganus_oreganus_cytochrome_b:0.00800,Crotalus_horridus_cytochrome_b:0.05866):0.04732,(Thamnophis_elegans_terrestris_cytochrome_b:0.00366,Thamnophis_atratus_cytochrome_b:0.00172):0.06255):0.00555,(Pituophis_catenifer_vertebralis_cytochrome_b:0.00552,Lampropeltis_getula_cytochrome_b:0.02035):0.05762,((Diadophis_punctatus_cytochrome_b:0.06486,Contia_tenuis_cytochrome_b:0.05342):0.01037,Hypsiglena_torquata_cytochrome_b:0.05346):0.00779);"
)
```

```{r}
library(ape)
data(bird.orders)
treewidget(bird.orders, height=400, width=800)
```

## packwidget
In many viral studies, taking a phylogeny and making clusters of closely related taxa (basically OTUs) is a big thing. Here is a basic interactive plot that packs circles with a given radius belonging to several different clusters, based on the work of @mbostock.
```{r}
r <- rpois(20,lambda=20)
cl <- rep(seq(1,2),each=10)
packwidget(r,cl,browser=FALSE)
```

## matrixcube
This is the beginnings of a viewer to visualize contacts between individuals at different times, a bit like in Cubix. This is based heavily on the 3D scatterplot example from rthreejs by @bwlewis.
```{r}
cf <- c("a", "b", "a", "c", "d")
ct <- c("b", "c", "c", "d", "b")
nodes <- sort(unique(c(cf,ct)))
x <- match(cf,nodes)
y <- match(ct,nodes)
z <- c(1, 2, 3, 4, 5)
matrixcube(x,y,z,axis.labels=c("Ego","Alter","Onset"),tick.labels=list(nodes,nodes,sort(unique(z))))
```

