---
title: "Learn HTMLWidget"
author: "Ko, Kuei Yueh"
date: "Thursday, July 16, 2015"
output: html_document
---
```{r}
setwd("C:\\Users\\clinton\\GitHub\\CodeLearn\\LearnR")
```

# Install and Import the libraries
```{r}
#library(devtools)
# time seires charting
#install_github("rstudio/dygraphs")
# DataTables
#install_github("rstudio/DT")
# Leaflet maps
#install_github("rstudio/leaflet")
# network graphs by D3
#install.packages("networkD3")
# 3D scatterplots and globe (three.js)
#install_github("bwlewis/rthreejs")
# 3D scatterplots, lines and bars (MetricsGraphics.js)
#install_github("hrbrmstr/metricsgraphics")
# Interactive Heatmaps (d3heatmap)
#install_github("rstudio/d3heatmap")
# DigrammeR
#install_github("rich-iannone/DiagrammeR")
```

Note:  
The below tutorial is from
http://www.htmlwidgets.org/
https://www.youtube.com/watch?v=dV4gtARPvu8

# Leaflet
http://rstudio.github.io/leaflet/

Leaflet is a JavaScript library for creating dynamic maps that support panning and zooming along with various annotations like markers, polygons, and popups.

Tutorial from htmlwidgets.org
```{r}
library(leaflet)
orstationc <- read.csv("C:/Users/clinton/GitHub/CodeLearn/LearnR/data/orstationc.csv")

pal <- colorQuantile("YlOrRd", NULL, n=8)
leaflet(orstationc) %>%
    addTiles %>%
    addCircleMarkers(color = ~pal(tann))
```

# Dygraphs
http://rstudio.github.io/dygraphs

Dygraphs provides rich facilities for charting time-series data in R and includes support for many interactive features including series/point highlighting, zooming, and panning.

Tutorial from htmlwidgets.org
```{r}
library(dygraphs)
dygraph(nhtemp, main="New Haven Temperatures") %>%
    dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

# MetricsGraphics
http://hrbrmstr.github.io/metricsgraphics/

MetricsGraphics enables easy creation of D3 scatterplots, line charts, and histograms.

Tutorial from htmlwidgets.org
```{r}
library(metricsgraphics)
mjs_plot(mtcars, x=wt, y=mpg) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")
```

# NetworkD3
http://christophergandrud.github.io/networkD3/

networkD3 provides tools for creating D3 JavaScript network graphs from R.

Tutorial from htmlwidgets.org
```{r}
library(networkD3)
data(MisLinks, MisNodes)
forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.4)
```

# d3Heatmap
https://github.com/rstudio/d3heatmap

Interactive heatmaps with D3 including support for row/column highlighting and zooming.

Tutorial from htmlwidgets.org
```{r}
library(d3heatmap)
d3heatmap(mtcars, scale="column", colors="Blues")
```

# DataTables
http://rstudio.github.io/DT/

DataTables displays R matrices or data frames as interactive HTML tables that support filtering, pagination, and sorting.

Tutorial from htmlwidgets.org
```{r}
library(DT)
datatable(iris, options = list(pageLength = 5))
```

# threejs
https://github.com/bwlewis/rthreejs

threejs includes a 3D scatterplot and 3D globe (you can directly manipulate the scatterplot below with the mouse).

Tutorial from htmlwidgets.org
```{r}
library(threejs)
z <- seq(-10, 10, 0.01)
x <- cos(z)
y <- sin(z)
scatterplot3js(x,y,z, color=rainbow(length(z)))
```

# DiagrammeR
http://rich-iannone.github.io/DiagrammeR/

A tool for creating diagrams and flowcharts using Graphviz and Mermaid.

Tutorial from htmlwidgets.org
```{r}
library(DiagrammeR)
grViz("
  digraph {
    layout = twopi
    node [shape = circle]
    A -> {B C D} 
  }")
```

