---
title: "Learn Plotly: Giving a Taste"
author: "Ko, Kuei Yueh"
date: "Saturday, September 12, 2015"
output: html_document
---

The practice is from the following website  
http://www.r-bloggers.com/3d-plots-with-ggplot2-and-plotly/

# install the package
```{r}
#install.packages("Rcpp")
#install.packages("devtools")
#library(devtools)
#install_github("ropensci/plotly")  # plotly is part of ropensci
```

# Get Started
To connect to the plotly web platform, the username and key must be first set.  
(https://plot.ly/r/getting-started/)
```{r}
library(plotly)

# set the username and key in R
Sys.setenv("plotly_username"="clintko")
Sys.setenv("plotly_api_key"="wzlp2ee5mi")
```

# Start plotting
The figure below is from the link  
http://blog.plot.ly/post/106630252117/plot-with-ggplot2-and-plotly-within-knitr-reports
```{r}
p <- plot_ly(economics, x = date, y = uempmed, 
             filename="LearnPlotly/economics_dat_uempmed")
p
```

#----------------------------------------------------------
The figures in this section is from the link: "https://plot.ly/r/"

create the data
```{r}
set.seed(100)
d <- diamonds[sample(nrow(diamonds), 1000), ]
head(d)
```

plot the carat vs price
```{r}
plot_ly(d, x = carat, y = price, 
        text = paste("Clarity: ", clarity),
        mode = "markers", 
        color = carat, 
        size = carat,
        filename="LearnPlotly/diamonds_carat_vs_price")
```

replot the scatterplot with ggplot
```{r}
qplot(carat, price, data=d, colour=clarity)
ggplotly(filename="LearnPlotly/diamonds_carat_vs_price_ggplot")
```

replot using ggplot function. Stratefy the data using the colnames "cut" and plot the scatter plot using geomPoint. Fianlly, fit each data with loess in geomSmooth.
```{r}
p <- ggplot(data = d, aes(x = carat, y = price)) +
  geom_point(aes(text = paste("Clarity:", clarity)), size = 4) +
  geom_smooth(aes(colour = cut, fill = cut)) + facet_wrap(~ cut)

(gg <- ggplotly(p, filename="LearnPlotly/diamonds_curveFitting"))
```

the plotly support the 3D plot as well (which is the main reason I learn plotly)
```{r}
plot_ly(d, x=carat, y=price, z=depth, 
        type = "scatter3d", mode = "markers", 
        marker = list(size = 2), 
        filename="LearnPlotly/diamonds_3DPlot_carat_price_depth")
```


